Performance of two different ontological and data access methods

### 2014-07-02 cold2014/magnus.tex
subsection experiments
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@706 43ef1190-90fd-4dd7-a769-d0121f02b80e

\subsection{Experiment with three factors}
The first experiment that was done, was the experiment with only three
factors. The goal here was to see how the ontologies behave with a
small amount of querying and factors.  Table \ref{facandlevExp1} show
the different factors and their levels. We chose to start with a small
experiment using only simple factors which had few possible
values. That is why we chose transmission and fuel type as factors as
well as the ontology factor. With these three factors we got $2^3$
number of runs.  This meant that we had 8 runs to execute against our
application. This experiment was also replicated four times to get the
most accurate results. The experiment was small so the replication did
not cause any issues.  The results were calculated by finding the
average response time for each run between all four replications.

@@

\subsection{Experiment with ten factors}
The last experiment we did was a ten factor experiment. The goal here
was to see which factors would be deemed significant in a complete car
model search. With the results from this experiment we could discuss
several aspects of using the different ontologies.  Here we added 4
new factors in comparison to the six factor experiment. The newly
added factors were \emph{weight}, \emph{nr. of doors},
\emph{nr. of seats} and \emph{fuel consumption}. We used the same
approach here as in the previous experiments to calculate the average
value for the new factors. The average values can be seen in the
planning matrix in appendix \ref{appB}. With 10 factors the amount of
runs were $2^{10}$. This results in 1024 number of runs which made the
planning matrix too big to show here.

@@

We recommend that both alternatives should be used together for
representing complex products. % TODO: Spiser ikke helt denne
Using only CO there should be done some alterations to reduce
bottleneck issues. We would not recommend using the data represented
by Renault today, because it is not robust enough and has a too long
response time for bigger operations.  It is also important that the
data representation is done properly to ensure readability for both
humans and machines.  Today, the Renault representation is neither
very understandable for machine nor for human. This is reflected in
the discussion about the application development.  The results from
the experiments done against the application shows that the current
representation is not feasible with today's performance standards and
user expectations. This is the area where most of the workload is on
the use of the ontologies.  This means that for a third party user VSO
in conjunction with COO are the best solution to work with product
models. However, this approach is limited to only represent data about
car models. It would also require some reasoning to be effective, but
with the technology and knowledge today that should not be a to
difficult issue. For other users wanting to represent constraints
between product models, CO can be a viable options if the proper
alterations is done. It can be very useful when working with several
data domains, but then the compatibility issue has to be solved.  One
solution could be to use the compatibility approach that COO
offers. This would result in adding several new triples in each
lexicon to avoid the configuration traversals.  It would also add the
need for reasoning to be effective, which is opposite of what Renault
proposed. If the compatibility issue is solved we believe the
performance can be feasible for almost every scenario. It will demand
more from the third party user, but can utilize more than just data
about car models. To make it even more effective, there should be done
several other alterations to both the representation and the ontology
as discussed in Section \ref{Alterations}.

>>>>

We ran several experiments, both randomized and unrandomized
experiments, with 3, 4, 6 and 10 factors. Here, we report on the
unrandomized 3 and 10 full factorial experiments:

\subsubsection{Experiment with three factors}
The first experiment that was done, was the experiment with only three
factors. The goal here was to see how the ontologies behave with a
small amount of querying and factors.  Table \ref{facandlevExp1} show
the different factors and their levels. We chose to start with a small
experiment using only simple factors which had few possible
values. That is why we chose transmission and fuel type as factors as
well as the ontology factor. With these three factors we got $2^3$
number of runs.  This meant that we had 8 runs to execute against our
application. This experiment was also replicated four times to get the
most accurate results. The experiment was small so the replication did
not cause any issues.  The results were calculated by finding the
average response time for each run between all four replications.

@@

\subsubsection{Experiment with ten factors}
The last experiment we did was a ten factor experiment. The goal here
was to see which factors would be deemed significant in a complete car
model search. With the results from this experiment we could discuss
several aspects of using the different ontologies.  Here we added 4
new factors in comparison to the six factor experiment. The newly
added factors were \emph{weight}, \emph{nr. of doors},
\emph{nr. of seats} and \emph{fuel consumption}. We used the same
approach here as in the previous experiments to calculate the average
value for the new factors. The average values can be seen in the
planning matrix in appendix \ref{appB}. With 10 factors the amount of
runs were $2^{10}$. This results in 1024 number of runs which made the
planning matrix too big to show here.

@@

We recommend that both alternatives should be used together for
representing complex products. 
Using only CO there should be done some alterations to reduce
bottleneck issues. We would not recommend using the data represented
by Renault today, because it is not robust enough and has a too long
response time for bigger operations.  It is also important that the
data representation is done properly to ensure readability for both
humans and machines.  Today, the Renault representation is neither
very understandable for machine nor for human. This is reflected in
the discussion about the application development.  The results from
the experiments done against the application shows that the current
representation is not feasible with today's performance standards and
user expectations. This is the area where most of the workload is on
the use of the ontologies.  This means that for a third party user VSO
in conjunction with COO are the best solution to work with product
models. However, this approach is limited to only represent data about
car models. It would also require some reasoning to be effective, but
with the technology and knowledge today that should not be a to
difficult issue. For other users wanting to represent constraints
between product models, CO can be a viable options if the proper
alterations is done. It can be very useful when working with several
data domains, but then the compatibility issue has to be solved.  One
solution could be to use the compatibility approach that COO
offers. This would result in adding several new triples in each
lexicon to avoid the configuration traversals.  It would also add the
need for reasoning to be effective, which is opposite of what Renault
proposed. If the compatibility issue is solved we believe the
performance can be feasible for almost every scenario. It will demand
more from the third party user, but can utilize more than just data
about car models. To make it even more effective, there should be done
several other alterations to both the representation and the ontology
as discussed in Section \ref{Alterations}.
